---
title: "Mexico Choropleths"
output: html_document
fig_width: 8
fig_height: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# mxmaps

[![Travis-CI Build Status](https://travis-ci.org/diegovalle/mxmaps.svg?branch=master)](https://travis-ci.org/diegovalle/mxmaps) 
[![Coverage Status](https://coveralls.io/repos/github/diegovalle/mxmaps/badge.svg?branch=master)](https://coveralls.io/github/diegovalle/mxmaps?branch=master)

This R package can be used to easily create maps of Mexico at both the state and municipio levels. It is based on the [choroplethr](https://cran.r-project.org/web/packages/choroplethr/index.html) package and includes functions to create interactive maps using the leaflet package, map INEGI data using their API with the [inegiR](https://github.com/Eflores89/inegiR) package, and format strings so they match the INEGI state and municipio codes.


## Installation

For the moment this package is only available from github. For the development version:

```r
if (!require("devtools")) {
    install.packages("devtools")
}
devtools::install_github("diegovalle/mxmaps")
```

## Quick Example

```{r}
library("mxmaps")

df_mxstate$value <- df_mxstate$pop
mxstate_choropleth(df_mxstate,
                   title = "Total population, by state") 
# What the data looks like
head(df_mxstate[,c("region", "value")])
```

##Basic Usage

The two basic functions of the package are ```mxstate_choropleth``` and ```mxmunicipio_choropleth```. The data.frame that you provide to the plotting functions must have one column named ```region``` with the INEGI codes of the states or municipios and one column named ```value``` with the values to plot. You can specify a title with the ```title``` parameter and the number of buckets with the ```num_color``` parameter. If you want a continuous scale you can set the ```num_colors``` parameter equal to 1.

```{r}
library("mxmaps")

df_mxstate$value <-  df_mxstate$indigenous / df_mxstate$pop 
mxstate_choropleth(df_mxstate, 
                   num_colors = 1,
                   title = "Percentage of the population that self-identifies as indigenous")
```

## Municipios

You can also plot Mexican municipios (similar to counties) with a continuous color scale set with num_colors = 1

```{r}
data("df_mxmunicipio")
df_mxmunicipio$value <-  df_mxmunicipio$indigenous / df_mxmunicipio$pop 
mxmunicipio_choropleth(df_mxmunicipio, 
                       num_colors = 1,
                       title = "Percentage of the population that self-identifies as indigenous")

```

You can also subset the area to show in the choropleth by using the zoom parameter:

```{r}

mxmunicipio_choropleth(df_mxmunicipio, num_colors = 1,
                       zoom = subset(df_mxmunicipio, metro_area %in% c("Valle de México",
                                                                       "Puebla-Tlaxcala",
                                                                       "Cuernavaca",
                                                                       "Toluca"))$region,
                       title = "Percentage of the population that self-identifies as indigenous") 
```

You can use the ```show_states``` parameter to hide or show the state borders when making municipio choropleths.

```{r}
mxmunicipio_choropleth(df_mxmunicipio, num_colors = 1,
                       zoom = c(31001:31106,30002:30212),
                       title = "Percentage of the population that self-identifies\nas indigenous in Yucatán and Veracruz",
                       show_states = FALSE)
```
