---
title: "Data"
output: html_document
css: css/report_styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.dim = c(6, 6), 
                      dpi = 100, 
                      out.width = '100%')
```

The data.frame that you provide to the plotting functions must have one column named "region" and one column named "value". The entries for "region" must match the INEGI codes for states ("01", "02", etc) or municipios ("01001", "01002", etc) either as strings with or without a leading "0" or as numerics. The functions `str_mxstate` and `str_mxmunicipio` are provided to easily format codes to the INEGI specification. Also, some example data.frames, such as `df_mxstate_2020` and `df_mxmunicipio_2020`, are provided with demographic variables from the 2020 Census.

```{r}
library(mxmaps)
library(htmlTable)

data("df_mxstate_2020")
htmlTable(head(df_mxstate_2020), rnames = FALSE, col.rgroup = c("none", "#F7F7F7"))
data("df_mxmunicipio_2020")
htmlTable(head(df_mxmunicipio_2020), rnames = FALSE, col.rgroup = c("none", "#F7F7F7"))
```

## 2015 Conteo Intercensal

There are additional data.frames with data from other censuses or conteos 
intercensales

```{r}
data("df_mxstate_2015")
htmlTable(head(df_mxstate_2015), rnames = FALSE, col.rgroup = c("none", "#F7F7F7"))
data("df_mxmunicipio_2015")
htmlTable(head(df_mxmunicipio_2015), rnames = FALSE, col.rgroup = c("none", "#F7F7F7"))
```

## Older Censuses

```{r}
data("df_mxstate_1990_2010")
htmlTable(head(df_mxstate_1990_2010), rnames = FALSE, col.rgroup = c("none", "#F7F7F7"))
data("df_mxmunicipio_1990_2010")
htmlTable(head(df_mxmunicipio_1990_2010), rnames = FALSE, col.rgroup = c("none", "#F7F7F7"))
```

State names and abbreviations come in two versions since normal people always use _Veracruz_ instead of its full official name of _Veracruz de Ignacio de la Llave_, and it's kind of awkward to use the Chiapas official abbreviation of _Chis._

```{r message=FALSE,warning=FALSE}
library("geojsonio")

# Convert the topoJSON to spatial object
data(mxstate.topoJSON)
tmpdir <- tempdir()
# have to use RJSONIO or else the topojson isn't valid
write(RJSONIO::toJSON(mxstate.topoJSON), file.path(tmpdir, "state.topojson"))
states <- topojson_read(file.path(tmpdir, "state.topojson"))
```

## String conversion

```{r}
str_mxstate(c(32, 5))
str_mxmunicipio(c(32, 5), c(9, 18))
str_mxmunicipio(c(32001, 5009))
```
