---
title: "Leaflet States"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.dim = c(6, 6), 
                      dpi = 100, 
                      out.width = '100%')
```

This document provides examples of how to create interactive choropleth maps of Mexican states using the `mxmaps` and `leaflet` packages. The examples illustrate how to build maps with both continuous and categorical data, and how to customize the map legends.

```{r}
library(mxmaps)
library(leaflet) # for colorNumeric
library(scales) # needed for comma

df_mxstate_2020$value <- df_mxstate_2020$afromexican / df_mxstate_2020$pop
pal <- colorNumeric("Blues", domain = df_mxstate_2020$value)
mxstate_leaflet(df_mxstate_2020,
                pal,
                ~ pal(value),
                ~ sprintf("State: %s<br/>Percent Afro-Mexican: %s",
                          state_name, comma(value))) %>%
  addLegend(position = "bottomright", 
            pal = pal, 
            values = df_mxstate_2020$value,
            title = "Percent<br>Afro-Mexican",
            labFormat = labelFormat(suffix = "%",
                                    transform = function(x) {100 * x})) %>%
  addProviderTiles("CartoDB.Positron")
```


You can also create maps with categorical variables by first converting the 
variables to integers and using the `colors` and `labels` parameters to `addLegend` to assign the colors to each level in the variable.


```{r}
library(mxmaps)
library(leaflet) # for colorBin
library(scales) # needed for comma

df_mxstate_2020$value2 <- factor(sample(c("a", "b", "c", "d", "e"), 32, replace = TRUE))
df_mxstate_2020$value <- as.integer(df_mxstate_2020$value2)

pal <- colorBin(rainbow(5), domain = as.integer(factor(df_mxstate_2020$value2)))
mxstate_leaflet(df_mxstate_2020,
                pal,
                ~ pal(value),
                ~ sprintf("State: %s<br/>Percent Afro-Mexican: %s",
                          state_name, comma(value))) %>%
  addLegend(position = "bottomright", 
            colors =c(rainbow(5)),
            labels= c("a", "b","c","d", "e"),
            title = "categoría"
  ) %>%
  addProviderTiles("CartoDB.Positron")

```