---
title: "Municipios"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.dim = c(6, 4.6),
  dpi = 200,
  out.width = "100%"
)
```

This document provides examples of how to create choropleth maps of Mexican municipios using the `mxmaps` package. The examples showcase various customizations, such as using continuous and categorical data, zooming into specific regions, removing municipio borders, and adding labels to the map.

### Continuous Color Scale

The following example demonstrates how to create a choropleth map of all Mexican municipios with a continuous color scale. The map shows the percentage of the population that speaks an indigenous language.

```{r message=FALSE,warning=FALSE}
library(mxmaps)
library(viridis)
library(scales)

df_mxmunicipio_2020$value <- df_mxmunicipio_2020$indigenous_language /
  df_mxmunicipio_2020$pop
gg <- MXMunicipioChoropleth$new(df_mxmunicipio_2020)
gg$title <- "Percentage of the population that speaks\nan indigenous language"
gg$set_num_colors(1)
gg$ggplot_scale <- scale_fill_viridis("percent",
  labels = percent
)
gg$render()
```

### Categorical Data and Zooming

This example shows how to create a map with categorical data. The map is zoomed in to show only the municipios in the states of Mexico and Mexico City. The `theme_void()` function is used to create a clean layout without axes or gridlines.

```{r message=FALSE,warning=FALSE}
library("ggplot2")

df_mxmunicipio_2020$value <- as.factor(sample(c(NA, letters[1:6]),
  nrow(df_mxmunicipio_2020),
  replace = TRUE
))
gg <- MXMunicipioChoropleth$new(df_mxmunicipio_2020)
gg$title <- "Municipios a-f"
gg$set_num_colors(6)
gg$set_zoom(subset(df_mxmunicipio_2020, state_name %in% c(
  "Ciudad de México",
  "México"
))$region)
gg$ggplot_scale <- scale_fill_brewer("type",
  type = "qual", palette = 2,
  na.value = "gray"
)
p <- gg$render()
p + theme_void()
```

### Removing Municipio Borders

The `mxmunicipio_choropleth` function returns a `ggplot` object, which can be further customized. In this example, the borders of the municipios are removed by directly modifying the `ggplot` object.

```{r message=FALSE,warning=FALSE}
library("scales")

df_mxmunicipio_2020$value <- df_mxmunicipio_2020$afromexican /
  df_mxmunicipio_2020$pop
p <- mxmunicipio_choropleth(df_mxmunicipio_2020,
  title = "Percentage of population that identifies as Afromexican",
  legend = "percent\nAfromexican",
  num_colors = 1
)
p[["layers"]][[1]][["aes_params"]][["colour"]] <- "transparent"
p
```

### Adding Labels to the Map

This example demonstrates how to add labels to the map to identify important municipios. The map is zoomed in to the state of Chihuahua, and labels are added for all municipios with a population greater than 100,000. The `ggrepel` package is used to prevent the labels from overlapping.

```{r}
library("ggrepel")

df_mxmunicipio_2020$value <- df_mxmunicipio_2020$indigenous_language /
  df_mxmunicipio_2020$pop * 100
chih <- subset(df_mxmunicipio_2020, state_name == "Chihuahua")
p <- mxmunicipio_choropleth(df_mxmunicipio_2020,
  num_colors = 1,
  zoom = chih$region,
  title = "Percentage of the population that self-identifies\nas indigenous in Chihuahua",
  show_states = FALSE,
  legend = "%"
)

labels <- chih
labels$group <- NA
labels <- subset(labels, pop > 1e05)
p +
  geom_text_repel(
    data = labels,
    aes(long, lat, label = municipio_name),
    nudge_x = 0.1,
    nudge_y = 0.7
  ) +
  geom_point(
    data = labels,
    aes(long, lat),
    color = "#d6604d",
    size = 1
  )
```

### Scale Bar

mxmaps can automatically add a scale bar to your map, making it easier to interpret distances at a glance. The scale bar can be customized through several parameters: `scale_bar_position` controls where it is placed on the map (for example, `"bl"` for bottom-left), `scale_bar_length` sets the total length represented by the scale bar in kilometers, `scale_bar_segments` defines how many divisions the bar is split into, `scale_bar_height` adjusts its visual height, while `scale_bar_color` and `scale_bar_text_color` control the colors of the bar itself and its labels. For example, `scale_bar_position = "bl", scale_bar_length = 1000`, `scale_bar_segments = 4`, `scale_bar_height = 0.5`, `scale_bar_color = "black"`, `scale_bar_text_color = "black"` will place a 1000 km scale bar with four segments in the bottom-left corner of the map using black styling.

```{r}
df_mxmunicipio_2020$value <- df_mxmunicipio_2020$indigenous_language /
  df_mxmunicipio_2020$pop * 100
mxmunicipio_choropleth(df_mxmunicipio_2020,
  num_colors = 1,
  title = "Percentage of the population that speaks\nan indigenous language",
  scale_bar = TRUE,
  scale_bar_position = "bl",
  scale_bar_length = 1000,
  scale_bar_segments = 4,
  background_color = "#e3dac9",
  title_color = "#635147",
  municipio_border_color = "saddlebrown",
  municipio_border_size = 0.01,
  state_border_color = "#635147",
  state_border_size = .5,
  legend = "%"
)
```
